---
title: "PA1_template"
author: "Roman Braun"
date: "04 12 2017"
output: 
        md_document:
        variant: markdown_github
---

### Downloading and opening data

This document will download and analyze the [Activity Monitoring Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip).
First, set working directory and check for existence of the .zip file. Second, 
unzip the file to get access to activity.csv. Third, open csv and look at the data
set.

```{r download & open, echo = TRUE}
wd <- "/Users/romanbraun/Dropbox (Privat)/Coursera/Data Science/5_reproducible_research/W2 - assignment"
setwd(wd)
if (file.exists("repdata_data_activity.zip")) {
        unzip("repdata_data_activity.zip")
} else {
        download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", 
        "repdata_data_activity.zip")
        unzip("repdata_data_activity.zip")
}
df <- read.table("activity.csv", sep=",", header = TRUE)  
#head(df)
#dim(df)
#as.Date(df$date[2])
df$date <- as.Date(df$date)
```

### Calculating  mean number of steps per day

You can also embed plots, for example:

```{r steps per day plot, echo=TRUE}
steps_sum <- aggregate(steps ~ date, data = df, sum) 
#dim(steps_sum)
#steps_sum
# Print steps per day as list
#steps_sum
library(ggplot2)
qplot(x = steps_sum$steps, xlab = "steps", binwidth = 500)

```

Figure 1: Distribution of sum of steps in dataset

Mean steps per day:
```{r}
mean(steps_sum$steps)
```
Median steps per day:
```{r}
median(steps_sum$steps)
```

### What is the daily pattern?
```{r}
#head(df)
#class(df)
steps_int <- aggregate(df, by=list(df$interval), FUN=mean, na.rm=TRUE) 
#steps_int
ggplot(steps_int, aes(x = interval, y = steps)) +
        geom_line(stat="identity")

```

Figure 2: Daily pattern of steps throughout data set

Maximum of steps
```{r}
steps_int[which.max(steps_int$steps),c(1,2)]
```

### Imputing missing values


### Removing NA values
Replace NA values by '0'.

```{r}
dfnew <- df
dfnew$steps[is.na(dfnew$steps)] <- 0
sum(is.na(dfnew$steps))

```

Histogram of total number of steps including imputed values: 
```{r}

steps_sum <- aggregate(steps ~ date, data = dfnew, sum)
library(ggplot2)
qplot(x = steps_sum$steps, xlab = "steps", binwidth = 500)
```

Figure 3: Distribution of number of steps after imputing missing values

Mean steps per day with imputed values:
```{r}
mean(steps_sum$steps)
```
Median steps per day with imputed values:
```{r}
median(steps_sum$steps)
```



### Are there differences in activity patterns between weekdays and weekends?
```{r}
df$weekday <- weekdays(as.Date(df$date))
mofr <- c("Montag", "Dienstag", "Mittwoch", "Donnerstag", "Freitag")
saso <- c("Samstag", "Sonntag")
df_week <- df[df$weekday == mofr,]
df_week <- subset(df, weekday %in% mofr)
df_wend <- subset(df, weekday %in% saso)
sel <- c("steps", "interval")
summary(df_week)

df_week <- df_week[,sel]
#head(steps_week)
steps_week <- aggregate(df_week, by=list(df_week$interval), FUN=mean, na.rm=TRUE) 
ggplot(steps_week, aes(x = interval, y = steps)) +
        geom_line(stat="identity")

```

Figure 4: Steps pattern on week days (Monday - Friday).


```{r}
df_wend <- df_wend[,sel]
steps_wend <- aggregate(df_wend, by=list(df_wend$interval), FUN=mean, na.rm=TRUE) 
#steps_int
ggplot(steps_wend, aes(x = interval, y = steps)) +
        geom_line(stat="identity")
```

Figure 5: Steps pattern on weekend (Saturday / Sunday).

### Create new dataset
```{r}
write.table(dfnew, "activity_new.csv", sep =";", row.names = FALSE)
steps_sum_new <- aggregate(steps ~ date, data = dfnew, sum) 
dim(steps_sum_new)
# Print steps per day as list
#steps_sum


library(ggplot2)

ggplot(steps_sum_new, aes(x=date, y=steps)) +
        geom_bar(stat = "identity")
```

### Conclusion
There's no difference in the histogram by imputing missing values, because NA was replace by '0' which makes sense in this context. However, the imputing of "0" had an effect on the calculation of the mean and median number of steps per day.
No data available means that there were no steps performed.
Analysis of steps pattern on weekdays and weekends shows substantial differences.