---
title: "PA1_template"
author: "Roman Braun"
date: "26 10 2017"
output: html_document
---

### Downloading and opening data

This document will download and analyze the [Activity Monitoring Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip).
First, set working directory and check for existence of the .zip file. Second, 
unzip the file to get access to activity.csv. Third, open csv and look at the data
set.

```{r download & open, echo = TRUE}
wd <- "/Users/romanbraun/Dropbox (Privat)/Coursera/Data Science/5_reproducible_research/W2 - assignment"
setwd(wd)
if (file.exists("repdata_data_activity.zip")) {
        unzip("repdata_data_activity.zip")
} else {
        download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", 
        "repdata_data_activity.zip")
        unzip("repdata_data_activity.zip")
}
df <- read.table("activity.csv", sep=",", header = TRUE)  
#head(df)
#dim(df)
#as.Date(df$date[2])
df$date <- as.Date(df$date)
```

### Calculating  mean number of steps per day

You can also embed plots, for example:

```{r steps per day plot, echo=TRUE}
steps_sum <- aggregate(steps ~ date, data = df, sum) 
dim(steps_sum)
# Print steps per day as list
#steps_sum


library(ggplot2)

ggplot(steps_sum, aes(x=date, y=steps)) +
        geom_bar(stat = "identity")
        
```

Mean steps per day:
```{r}
mean(steps_sum$steps)
```
Median steps per day:
```{r}
median(steps_sum$steps)
```

### What is the daily pattern?
```{r}
head(df)

ggplot(df, aes(x = interval, y = steps)) +
        geom_line(stat="identity")

steps_max <- aggregate(steps ~ interval, data = df, sum) 
max(steps_max$steps)
steps_max[steps_max$steps == 10927,]
```

### Imputing missing values
```{r}
sum(is.na(df$steps))
```


### Removing NA values
Replace NA values by '0'.

```{r}
dfnew <- df
dfnew$steps[is.na(dfnew$steps)] <- 0
sum(is.na(dfnew$steps))

```

### Create new dataset
```{r}
write.table(dfnew, "activity_new.csv", sep =";", row.names = FALSE)
steps_sum_new <- aggregate(steps ~ date, data = dfnew, sum) 
dim(steps_sum_new)
# Print steps per day as list
#steps_sum


library(ggplot2)

ggplot(steps_sum_new, aes(x=date, y=steps)) +
        geom_bar(stat = "identity")
```

### Conclusion
There's no difference by imputing missing values, because NA was replace by '0' which makes sense in this context.
No data available means that there were no steps performed.